An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-28 draft.tex
Update draft.tex
Experimental config

Figure \ref{fig:experiment_flow} shows the workflow users can follow to run
experiments in Evo\-Swarm. The first step is to develop the search strategy
algorithms that are going to be used by the multi-population algorithm.
Algorithm developers can upload the container definition or images directly to
DockerHub. Another option is to upload the image definition (Dockerfile) to
GitHub and define a trigger to re-upload the image to DockerHub.  Docker Compose
configuration files can reference these containers. When a user needs to deploy
the application, he or she needs to edit the {\tt docker-compose.yml} file to specify the
type and number of containers needed for the application.  The {\tt docker-compose}
application is responsible for starting all the containers or downloading new
versions if it is necessary. After starting the application, users can push
Experiment Definition files and start monitoring the execution in the terminal.
After the experiment is finished, the user must execute another script using the
experiment id to generate a collection of files containing all the data
generated by the experiment. The script generates these files in a JSON format.
Researchers can use these files to analyze and plot experiment data.  Python
scripts are included in the repository to plot the running times for the
experiments, and all other plots used in this paper.  There is also a script to
generate the files needed by the COCO framework \cite{hansen2016coco}, which can generate standard
comparisons against other methods.

@@

Next, we show the parameters for each algorithm, in Table \ref{tab:GAparams} 
for the GA and in Table \ref{tab:PSOparams} for PSO. We obtained these parameters 
following the same method as in a previous work \cite{garcia2017benchmarking}. 
To obtain the parameters, we tested first on the
Rastrigin separable function with five dimensions. After about fifteen
experiments, the most challenging targets were achieved for this particular
function. We tested again with functions one to three, and after obtaining
favorable results, the PSO and GA algorithm parameters were set. In the GA, we
randomly set the mutation and crossover probabilities to have more
heterogeneous workers; in these parameters, the range of values is specified.

@@

% Experiment configuration

\begin{table}
    \small
    \caption{ Experiment configuration example 
    }
    \label{tab:params}
    \centering
    \small
    \begin{tabular}{|l|l|l|l|l|}
      \hline
      \multicolumn{3}{|l|}{Parameter}                   & Type             & Example         \\ \hline
      \multicolumn{3}{|l|}{Worker Containers (Docker Compose)}        & \texttt{int}     & \texttt{8} \\ \hline
      \multicolumn{3}{|l|}{GA-PSO Ratio}                 & \texttt{decimal} & \texttt{0.5}    \\  \hline
      \multicolumn{3}{|l|}{Benchmark Functions}          & \texttt{list}    & \texttt{[1, 2, 3, 4, 5]} \\ \hline
      \multicolumn{3}{|l|}{Dimensions}                   & \texttt{list}    & \texttt{[10, 20, 40]}        \\ \hline
      \multicolumn{3}{|l|}{Crossover Probability Range}  & \texttt{list}    & \texttt{[0.2, 0.6]}      \\ \hline
      \multicolumn{3}{|l|}{Mutation  Probability Range} & \texttt{list}    & \texttt{[0.1, 0.3]}      \\ \hline
      \multicolumn{5}{|c|}{Dimensions}                                                      \\ \hline  
      Dimension               & Generations & Population Size & Populations  &     Iterations    \\ \hline
              10              & 50      & 140                 &      5                 & 30                \\ \hline
              20              & 66      & 200                 &      5                 & 30               \\ \hline
              40              & 80      & 250                 &      5                 & 40                \\ \hline
    \end{tabular}
\end{table}

>>>>

Figure \ref{fig:experiment_flow} shows the workflow users can follow to run
experiments in Evo\-Swarm. The first step is to develop the search strategy
algorithms that are going to be used by the multi-population algorithm.
Algorithm developers can upload the container definition or images directly to
DockerHub. Another option is to upload the image definition (Dockerfile) to
GitHub and define a trigger to re-upload the image to DockerHub.  Docker Compose
configuration files can reference these containers. When a user needs to deploy
the application, he or she needs to edit the Fi file to specify the
type and number of containers needed for the application.  The {\tt docker-compose}
application is responsible for starting all the containers or downloading new
versions if it is necessary. After starting the application, users can push
Experiment Definition files and start monitoring the execution in the terminal.
After the experiment is finished, the user must execute another script using the
experiment id to generate a collection of files containing all the data
generated by the experiment. The script generates these files in a JSON format.
Researchers can use these files to analyze and plot experiment data.  Python
scripts are included in the repository to plot the running times for the
experiments, and all other plots used in this paper.  There is also a script to
generate the files needed by the COCO framework \cite{hansen2016coco}, which can generate standard
comparisons against other methods.

@@

Next, we show the parameters for each algorithm, in Table \ref{tab:GAparams} for
the GA and in Table \ref{tab:PSOparams} for PSO. We obtained these parameters
following the same method as in a previous work \cite{garcia2017benchmarking}.
To obtain the parameters, we tested first on the Rastrigin separable function
with five dimensions. After about fifteen experiments, the most challenging
targets were achieved for this particular function. We tested again with
functions one to three, and after obtaining favorable results, the PSO and GA
algorithm parameters were set. In the GA, we randomly set the mutation and
crossover probabilities to have more heterogeneous workers; in these parameters,
the range of values is specified. We did not change these parameters during the
experiments, and only the population size and number of generations were
provided as parameters.

@@

% Experiment conf

The EvoSwarm application uses configuration files to run noiseless BOBB testbed
experiments. As a first step, the docker application must be deployed, using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided. Table \ref{tab:params} shows an example of these
parameters. The {\em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm for the search. With a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$ this will give us a PSO only algorithm, and finally, a
value of $1$ means that all populations will run the GA algorithm. Next, we
specify a list indicating which of the 24 benchmark functions will be tested.
In the example, the experiment will use the first five functions. The
{\em Instances} parameter, indicates how many instances of each function will be
executed, this parameter has 15 as the default value. This value is used
because is the standard for the BBOB benchmark \cite{hansen2016coco}. Then, in
the {\em Dimensions} parameter we define a list of the dimensions that will be
tested, we must select additional parameters for each dimension. According to
the maximum number of Function Evaluations (FEs) we mentioned earlier, we
define for each dimension: the number of populations that will be generated in
the setup, and for each population, the number of generations, and population
size. Finally, the {\em Iterations} parameter indicates how many complete loops
will be performed. All these parameters give us the maximum number of FEs that
will be performed. For instance, for $D = 2$, the maximum number of FEs is
$200,000$ which is the same as $40*50*10*10$.

\begin{table}
    \small
    \caption{ Experiment configuration example 
    }
    \label{tab:params}
    \centering
    \small
    \begin{tabular}{|l|l|l|l|l|}
      \hline
      \multicolumn{3}{|l|}{Parameter}                    & Type             & Example         \\ \hline
      \multicolumn{3}{|l|}{Worker Containers (Docker Compose)}        & \texttt{int}     & \texttt{8} \\ \hline
      \multicolumn{3}{|l|}{GA-PSO Ratio}                 & \texttt{decimal} & \texttt{0.5}    \\  \hline
      \multicolumn{3}{|l|}{Benchmark Functions}          & \texttt{list}    & \texttt{[1, 2, 3, 4, 5]} \\ \hline
      \multicolumn{3}{|l|}{Instances}                    & \texttt{integer}    & \texttt{15} \\ \hline
      \multicolumn{3}{|l|}{Dimensions}                   & \texttt{list}    & \texttt{[10, 20, 40]}        \\ \hline
      \multicolumn{3}{|l|}{Crossover Probability Range}  & \texttt{list}    & \texttt{[0.2, 0.6]}      \\ \hline
      \multicolumn{3}{|l|}{Mutation  Probability Range}  & \texttt{list}    & \texttt{[0.1, 0.3]}      \\ \hline
      \multicolumn{5}{|c|}{Dimensions}                                                      \\ \hline  
      Dimension               & Generations & Population Size & Populations  &     Iterations    \\ \hline
              10              & 50      & 140                 &      5                 & 30                \\ \hline
              20              & 66      & 200                 &      5                 & 30               \\ \hline
              40              & 80      & 250                 &      5                 & 40                \\ \hline
    \end{tabular}
\end{table}
